\sysfull

### 2018-04-16 paper/paging.tex
added section 5.4

\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data. During task execution, modifications to protected data do not affect their original memory locations, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{store}}.
\sys assumes that the write to a non-volatile word in memory is atomic.
%
%The non-volatile commit bit persists
%across power failures and \sys repeatedly attempts to commit until it succeeds,
%despite arbitrary power failures.
\subsection{Dynamic vs. Static Privatization}
\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
\TODO{pass by ref might not work the right way}
\begin{verbatim}

GLOBAL_(x)

foo(*y){ 
  *y++; 
}

TASK( ){ 
  ... ;  
  foo( &GV(x) ) ; 
}

\end{verbatim}
%\subsection{Managing Paging Overhead}
%
%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations. To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops. We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details). Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency. \sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
%
%\begin{figure}[t]
%	\centering
%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
%	\label{fig:dmaTimeEnergy}
%\end{figure}

>>>>

\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data. During task execution, modifications to protected data do not affect their original memory locations, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{store}}.
\sys assumes that the write to a non-volatile word in memory is atomic.
%
%The non-volatile commit bit persists
%across power failures and \sys repeatedly attempts to commit until it succeeds,
%despite arbitrary power failures.
\subsection{Dynamic Paging of \sys}

One design choice \sys made was to insert dynamic check for paging on every
possible problematic memory operation using {\tt RP} and {\tt WP} macros.
Since dynamic checks incur overhead on every read and write, it has a higher
overhead for accessing memory than a static approach,
e.g., Alpaca, where a system statically analyze the memory operation and try to insert only optimal
number of memory protection code (usually at the beginning of each task). 
%
However, \sys's dynamic paging has several benefits over static approach.
First, static approach has limitations, prohibiting certain types of programming pattern. For example,
in the presence of arbitrary pointer operation, function calls using function pointer, or an interrupt
within a task, the system cannot statically analyze the memory behavior. Thus, static
approaches either prohibits the programmer from using certain programming pattern~\cite{alpaca}, 
or will have to conservatively insert memory protection code that can be unnecessary.
Second, static approach cannot handle task coalescing, because with coalescing the scope to be
searched statically is not determined at compile time. 
By choosing to check every memory access dynamically, \sys supports programming pattern that
Alpaca does not support, such as arbitrary pointer operation, while also being able to
support task coalescing, which yields extra performance benefits.
%
%Moreover, if the programmer wants to optimize for high performance, it is possible for the
%programmer to place {\tt RP} and {\tt WP} only minimally by carefully examining the code.
%We argue that the

%\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
%\TODO{pass by ref might not work the right way}
%\begin{verbatim}
%
%GLOBAL_(x)
%
%foo(*y){ 
%  *y++; 
%}
%
%TASK( ){ 
%  ... ;  
%  foo( &GV(x) ) ; 
%}
%
%\end{verbatim}
%\subsection{Managing Paging Overhead}
%
%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations. To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops. We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details). Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency. \sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
%
%\begin{figure}[t]
%	\centering
%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
%	\label{fig:dmaTimeEnergy}
%\end{figure}
